Automatic Detection of Bad Programming Habits in Scratch: A Preliminary Study

### 2014-04-19 paperFIE_en.tex
Conclussions: add a short summary of the work done and plans for near future

\author{\IEEEauthorblockN{Jesús Moreno}
\IEEEauthorblockA{INTEF \& Programamos.es\\
Madrid, Spain\\
j.morenol@gmail.com}
\and
\IEEEauthorblockN{Gregorio Robles}
\IEEEauthorblockA{GSyC/LibreSoft\\Universidad Rey Juan Carlos\\
Madrid, Spain\\
grex@gsyc.urjc.es}}

@@

Scratch is being used both in extracurricular activities~\cite{maloney2008programming, kafai2010entering} or summer camps~\cite{adams2010scratching, franklin2013assessment}, as in elementary schools~\cite{wilson2012evaluation}, high schools~\cite{meerbaum2013learning} and even universities worldwide~\cite{wolz2009starting, malan2007scratch}. Its capacity to bring computer programming to children and teenagers with the aim of fostering computational thinking abilities and to promote ICT studies, and its success to teach both basic and advanced programming concepts (concepts such as user interaction, conditional statements, communication and synchronization, logical expressions and data representation are taught) has been documented (FIXME: reference required).

@@

FIXME: In this paper... short summary of what we have done and why

@@

In the near future, we plan to... FIXME... talk about next steps: i) more analysis, ii) more projects, iii) web-based tool that provides learning-oriented feedback in an automatic way.

>>>>

\author{\IEEEauthorblockN{Jesús Moreno}
\IEEEauthorblockA{INTEF \& Programamos.es\\
Madrid, Spain\\
jesus.moreno@programamos.es}
\and
\IEEEauthorblockN{Gregorio Robles}
\IEEEauthorblockA{GSyC/LibreSoft\\Universidad Rey Juan Carlos\\
Madrid, Spain\\
grex@gsyc.urjc.es}}

@@

Scratch is being used both in extracurricular activities~\cite{kafai2010entering} or summer camps~\cite{adams2010scratching}, as in elementary schools~\cite{wilson2012evaluation}, high schools~\cite{meerbaum2013learning} and even universities worldwide~\cite{malan2007scratch}. Its capacity to bring computer programming to children and teenagers with the aim of fostering computational thinking abilities and to promote ICT studies, and its success to teach both basic and advanced programming concepts (concepts such as user interaction, conditional statements, communication and synchronization, logical expressions and data representation are taught) has been documented~\cite{maloney2008programming, franklin2013assessment, wolz2009starting}.

@@

%FIXME: In this paper... short summary of what we have done and why

This paper shows the preliminary results of a study regarding two bad programming habits we have detected in our work as instructors with high school students learning to program with Scratch. In order to check if these bad habits, which have to do with object naming and code repetition, are also commonly found in the projects shared in the community repository, we downloaded 100 proyects and analyzed them with two plug-ins we developed for Hairball, detecting that most of the inspected proyects, 79\% and 62\%  respectively, fell into these issues.

@@

%In the near future, we plan to... FIXME... talk about next steps: i) more analysis, ii) more projects, iii) web-based tool that provides learning-oriented feedback in an automatic way.

In the near future we plan to extend the scope of our study developing new plug-ins to automate the detection of other issues, such as the degree of development of computational thinking demonstrated in a project, and analyzing a much larger number of projects. In this sense, the release of a research dataset with the first five years of data from the Scratch repository\footnote{https://sites.google.com/site/scratchdatameeting/} will be certainly a great help to our work. Finally, we are in the process of developing a website that will allow students to receive feedback from each of its programs with warnings of potential errors and tips to improve their code, and will offer teachers a tool to assist them in correcting their students proyects.
